---
title: "DataExploration"
author: "EllaJiang"
date: "1/30/2018"
output: html_document
---
```{r global_options, include=FALSE} 
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE) 
```

```{r load library}
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r get data}
dataset = readRDS('retail_one_percent.rds')
# dataset = dataset[sample(nrow(dataset),50000),]
options(scipen = 999) 
# add a new column: the sale date Y-M-D
dataset$date <- as.Date(with(dataset, paste(year_mo, mday,sep="-")), "%Y-%m-%d")
```

#### General Questions

* 1. Why some price are negative numebrs 
* 2. What are the meanings of variables: inv_strain, inv_productname? --- This is related to another problem: How can we decide the type of different marijuana product. This is important for all analysis by product type. I use invtype here for analysis.
* 3. As can be seen from the distributions below, there are some outliers, how can we deal with these outliers
* 4. There are some missing values in the dataset, such as the usable weight of some invtype, how can we deal with the missing vlaues. Also, there are some 0 in the usable weight column, which is not usable. How can we deal with this?
* 5. Can we Use CPI/Inflaction rate to adjust the price?
* 6. Do we need to predict the price in the future?


#### General Findings:

##### Frequency of different invtype across month and date in the dataset

```{r frequency}
# the occurance of each invtype by month
invtype.month = data.frame(with(dataset, table(year_mo, invtype)))
names(invtype.month)[3] = 'occurance'
inv.month.summary = ddply(invtype.month, 'year_mo', summarise, total = sum(occurance))
inv.month.summary$invtype = 'total'
names(inv.month.summary)[2] = 'occurance'
invtype.month = rbind(inv.month.summary,invtype.month)
ggplot(data = invtype.month, mapping = aes(x = year_mo, y = occurance, group = invtype, color = invtype)) + geom_line() + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) + labs(title = 'occurance of inventory type by month',y = 'occurance of transaction')

# the occurance of each invtype by day
invtype.day = data.frame(with(dataset, table(invtype,date)))
names(invtype.day)[3] = 'occurance'
inv.day.summary = ddply(invtype.day, 'date', summarise, total = sum(occurance))
inv.day.summary$invtype = 'total'
names(inv.day.summary)[2] = 'occurance'
invtype.day = rbind(inv.day.summary,invtype.day)
fre.day = ggplot(data = invtype.day, mapping = aes(x = date, y = occurance, group = invtype, color = invtype)) + geom_line() + labs(title = 'occurance of inventory type by day',y = 'occurance of transaction')
fre.day
```

* There is a clear increase in transactions of all inventory type.
* Usable Marijuana has many more transactions than other inventory type.

##### Analysis of Usable Weight

```{r usable weight}
# summary(dataset$usableweight)
# subset dataset
uw.data = select(dataset, sale_time, usableweight,invtype, year_mo,mday,date)

# the scatter plot of usableweigth over sale_time
ggplot(data = na.omit(uw.data), mapping = aes(x = sale_time , y = usableweight, color = invtype)) + geom_point() + labs(title = 'distribution of usable weight')

# the summary of usable weight for different invtype in different year_mo
uw.summary = ddply( uw.data, c('invtype','year_mo'), summarize, average.weight = mean(usableweight), max.weight = max(usableweight), min.weight = min(usableweight), median.weight = median(usableweight), std.weight = sd(usableweight))
uw.summary = arrange(uw.summary, invtype,year_mo)
uw.tendency = ggplot(uw.summary, mapping = aes(x=year_mo, y = average.weight, group = invtype, color = invtype)) + geom_line()  + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) + labs(title = 'average usable weight of inventory type by month',y = 'average usable weight')
uw.tendency

# the summary of usable weight for different invtype diffrent day
uw.summary = ddply( uw.data, c('invtype','date'), summarize, average.weight = mean(usableweight), max.weight = max(usableweight), min.weight = min(usableweight), median.weight = median(usableweight), std.weight = sd(usableweight))
uw.summary = arrange(uw.summary, invtype, date)
uw.day.tendency = ggplot(na.omit(uw.summary), mapping = aes(x=date, y = average.weight, group = invtype, color = invtype)) + geom_line()  + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) + labs(title = 'average usable weight of inventory type by day',y = 'average usable weigth')
uw.day.tendency
```

* There are many missing value, 0 and outliers in usableweight variable. How can we deal with them?
* I guess if we remove the outlier, the average usable weight of each invtype in each month could be stable.

##### Analysis on THC
```{r THC}
# subset dataset
THC.data = select(dataset, invtype,THC, usableweight, price_x,date,year_mo)
# distribution of THC and THC/usableweight
ggplot(data = THC.data, mapping = aes(x = date, y = THC)) + geom_point() + labs(title = 'Distribution of THC')
ggplot(data = THC.data, mapping = aes(x = date, y = THC/usableweight)) + geom_point() + labs(title = 'Distribution of THC/usableweight', ylab = 'THC/usableweight')

THC.subset = THC.data[apply(THC.data!=0, 1, all),]
THC.subset = mutate(THC.subset, THC_weight = THC/usableweight, THC_price = price_x/THC)
THC.summary = ddply(THC.subset, c("invtype","year_mo"), summarise, ave_THC_weight = mean(THC_weight,na.rm = TRUE), ave_THC_price = mean(THC_price, na.rm = TRUE) )
ggplot(data = na.omit(THC.summary), mapping = aes(x = year_mo, y = ave_THC_weight, group = invtype, color = invtype)) + geom_line() + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) + labs(title = 'average THC/usableWeight in different month')
ggplot(data = na.omit(THC.summary), mapping = aes(x= year_mo, y = ave_THC_price, group = invtype, color = invtype)) + geom_line() + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) + labs(title = 'average price_x/THC in different month')
```

##### Analysis on Price
```{r price}
price.data = select(dataset, price_x, usableweight, gram_price_x, invtype, year_mo, date)

ggplot(data = na.omit(price.data), mapping = aes(x=date, y = gram_price_x, color = invtype)) + geom_point() + labs(title = 'distribution of gram_price_x by date')

price.summary = ddply(price.data, c("year_mo","invtype"),summarise,gram_price_ave = mean(gram_price_x),gram_price_max = max(gram_price_x),gram_price_min = min(gram_price_x),gram_price_median = median(gram_price_x),gram_price_std = sd(gram_price_x))
ggplot(data = na.omit(price.summary), mapping = aes(x=year_mo, y = gram_price_ave, group = invtype, color = invtype)) + geom_line() + labs(title = 'average gram_price_x by month and product') + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1)) + geom_point()

price.summary2 = ddply(price.data, c("date","invtype"),summarise,gram_price_ave = mean(gram_price_x),gram_price_max = max(gram_price_x),gram_price_min = min(gram_price_x),gram_price_median = median(gram_price_x),gram_price_std = sd(gram_price_x))

ggplot(data = na.omit(price.summary2[price.summary2[,3]<150,]), mapping = aes(x=date, y = gram_price_ave, group = invtype, color = invtype)) + geom_point() + stat_smooth() + labs(title = 'average gram_price_x trendency by date of different product types') + theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1))

#price.data$rprice = round(price.data$gram_price_x, digits = 0)
#price.summary2 = ddply(na.omit(price.data), c('rprice'),summarise,count = n())
#price.summary2 = price.summary2[price.summary2[,1]<25,]
#price.summary2 = price.summary2[price.summary2[,1]> 0,]
#ggplot(data = price.summary2, mapping = aes(x = rprice, y = count)) + geom_line()
```