---
title: "price_invtype_exploration"
author: "EllaJiang"
date: "2/12/2018"
output: html_document
---
```{r global_options, include=FALSE} 
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE) 
```

```{r load library}
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

### load data and change some variable names

```{r get data}
# read data from the download folder
dataset = read.csv(file = "price_invtype_sample1.csv", header = TRUE)
# remove scientific notation
options(scipen = 999) 
# change the format of some variables
# lapply(dataset, class)
dataset$sale_date = as.Date(dataset$sale_date)
# generate year_mo and year_week variables
number1 = c("1","2","3","4","5","6","7","8","9")
number2 = c("01","02","03","04","05","06","07","08","09")
dataset$sale_month = mapvalues(as.character(dataset$sale_month), number1, number2)
dataset$sale_week = mapvalues(as.character(dataset$sale_week), number1, number2)
dataset = mutate(dataset, 
                 year_mo= paste(sale_year, sale_month, sep = "."),
                 year_week = paste(sale_year, sale_week, sep = "."))
```

### Transaction frequency of each inventory type across weeks

```{r transaction frequency across weeks}
dataset.count = dataset[which(dataset$invtype!=""),]
dataset.count = dataset[which(!is.na(dataset$sale_date)),]

# count the number of transaction of each inventory type in each week
count.week = ddply(dataset.count, 
                   c("year_week","invtype"), 
                   summarise, 
                   n.sold = n())

# plot the number of transaction of each inventory type in each week
ggplot(data = count.week, mapping = aes(x = year_week, y = n.sold, group = invtype, color = invtype)) +
  geom_line() + 
  labs(title = 'Transaction frequency of each inventory type cross weeks',y = 'transaction frequency') + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 6))
```

### Frequency of different inventory type across year months

```{r frequency in months}
# count the number of sells of each inventory type in each month
count.month = ddply(dataset.count, 
                    c("year_mo","invtype"), 
                    summarise, 
                    n.sold = n())

# plot the number of sells of each inventory type in each month (line chart)
ggplot(data = count.month, mapping = aes(x = year_mo , y = n.sold, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'Transaction frequency of inventory type cross months',y = 'transaction frequency') + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 6))
```

### Transaction Revenue of each inventory type across weeks
```{r revenue in weeks}
# remove null values from the dataset
dataset.revenue = dataset[which(dataset$invtype!=""),]
dataset.revenue = dataset.revenue[which(!is.na(dataset$price_adjust)),]

# sum the revenue for each inventory type in each week
revenue.week = ddply(dataset.revenue,
                     c("year_week","invtype"),
                     summarise,
                     revenue = sum(price_adjust))

# plot the revenue for each inventory type in each week
ggplot(data = revenue.week, mapping = aes(x = year_week, y = revenue, group = invtype, color = invtype, fill = invtype)) +
  geom_area()+
  labs(title = "Transaction revenue of inventory type cross weeks", y = "transaction revenue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 5))

```

### Transaction Revenue of each inventory type across months
```{r revenue in weeks2}
# sum the revenue for each inventory type in each week
revenue.month = ddply(dataset.revenue,
                     c("year_mo","invtype"),
                     summarise,
                     revenue = sum(price_adjust))

# plot the revenue for each inventory type in each week
ggplot(data = revenue.month, mapping = aes(x = year_mo, y = revenue, group = invtype, color = invtype, fill = invtype)) +
  geom_area()+
  labs(title = "Transaction revenue of inventory type cross months", y = "transaction revenue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
```

### Analysis on Price Using the first method: average(gram_price)

#### distribution of gram_price of each inventory type in different date
```{r average gram_price of each inventory type per week}
# remove price = 0 or usableweigth = 0 and missing values
dataset.price = subset(dataset, price_adjust!=0)
dataset.price = subset(dataset.price, usableweight_adjust != 0)

# calculate the gram_price of each transaction
dataset.price = mutate(dataset.price, gram_price = price_adjust/usableweight)

# one try of price distribution
# ----------------------------------------------------------------------
# # the distribution of gram_price of each inventory type
# dataset.price.nogroup = dataset.price[,c("sale_date","gram_price")]
# ggplot(data = dataset.price, mapping = aes(x=sale_date, y = gram_price)) + 
#   geom_point(data = dataset.price.nogroup, color= "grey") + 
#   geom_point(color = "black") + 
#   facet_wrap("invtype", nrow = 2, ncol = 2) + 
#   labs(title = 'distribution of gram_price by date')
#   
# # the distribution of gram_price of each inventory type (under 250)
# dataset.price.nogroup = dataset.price[,c("sale_date","gram_price")]
# ggplot(data = dataset.price, mapping = aes(x=sale_date, y = gram_price)) + 
#   geom_point(data = dataset.price.nogroup, color= "grey") + 
#   geom_point(color = "black") + 
#   facet_wrap("invtype", nrow = 2, ncol = 2) + 
#   labs(title = 'distribution of gram_price by date') +
#   ylim(0,250)
# 
# ggplot(data = dataset.price[which(dataset.price$invid == 28),], mapping = aes(x = price_adjust)) + 
#   geom_density()

# another try of price distribution 
# ------------------------------------------------
# subset dataset
inhalation = subset(dataset.price, invtype == "Marijuana Extract for Inhalation ")
infused = subset(dataset.price, invtype== "Marijuana Mix Infused ")
packaged = subset(dataset.price, invtype == "Marijuana Mix Packaged ")
usable = subset(dataset.price, invtype == "Usable Marijuana ")

# quantile plot of gram_price of each invtype
inha = ggplot(data = inhalation, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Marijuana Extract for Inhalation)')
infu = ggplot(data = infused, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Marijuana Mix Infused)')
pack = ggplot(data = packaged, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Marijuana Mix Packaged)')
usab = ggplot(data = usable, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Usable Marijuana)')

# put the graphs together
gridExtra::grid.arrange(inha, infu, pack, usab,ncol=2)
```

```{r}
# distribution of gram_price in each invtype
prob = c(0,0.05,0.1,0.5,0.9,0.95,1)
print("quantile distrition of Marijuana Extract for Inhalation")
data.frame(gram_price = quantile(inhalation$gram_price, prob))
print("quantile distrition of Marijuana Mix Infused")
data.frame(gram_price = quantile(infused$gram_price, prob))
print("quantile distrition of Marijuana Mix Packaged")
data.frame(gram_price = quantile(packaged$gram_price, prob))
print("quantile distrition of Usable Marijuana")
data.frame(gram_price = quantile(usable$gram_price, prob))

# 5% - 95% gram_price in each invtype
inhalation.new = subset(inhalation, 
                        gram_price < quantile(inhalation$gram_price, 0.95) & gram_price > quantile(inhalation$gram_price, 0.05))
infused.new = subset(infused, 
                        gram_price < quantile(infused$gram_price, 0.95) & gram_price > quantile(infused$gram_price, 0.05))
packaged.new = subset(packaged, 
                        gram_price < quantile(packaged$gram_price, 0.95) & gram_price > quantile(packaged$gram_price, 0.05))
usable.new = subset(usable, 
                        gram_price < quantile(usable$gram_price, 0.95) & gram_price > quantile(usable$gram_price, 0.05))

# quantile plot of gram_price of each invtype
inha.new = ggplot(data = inhalation.new, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Marijuana Extract for Inhalation)_New')
infu.new = ggplot(data = infused.new, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Marijuana Mix Infused)_New')
pack.new = ggplot(data = packaged.new, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)_New', 
       title = 'Quantile Plot \n gram_price (Marijuana Mix Packaged)')
usab.new = ggplot(data = usable.new, mapping = aes(sample = gram_price)) + 
  stat_qq(geom = 'point', distribution='qunif') + 
  labs(x='Quantiles', y='gram_price(dollars per gram)', 
       title = 'Quantile Plot \n gram_price (Usable Marijuana)_New')

# put the graphs together
gridExtra::grid.arrange(inha.new, infu.new, pack.new, usab.new,ncol=2)
```

#### distribution of avegram_price of each inventory type in each week

* with outliers (0%-100%)

```{r}
# calculate the average gram_price of each inventory type in each week
price.week = ddply(dataset.price, 
                   c("year_week","invtype"),
                   summarise,
                   gram_price_ave = mean(gram_price),
                   gram_price_max = max(gram_price),
                   gram_price_min = min(gram_price),
                   gram_price_median = median(gram_price),
                   gram_price_std = sd(gram_price))

# the distribution of average gram_price of each inventory type in each week
ggplot(data = price.week, mapping = aes(x=year_week, y = gram_price_ave, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'average gram_price by week and inventory type')+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))

# the distribution of median gram_price of each inventory type in each week
ggplot(data = price.week, mapping = aes(x=year_week, y = gram_price_median, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'median gram_price by week and inventory type')+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
```

* remove outliers (5%-95%)

```{r}
ggplot(data = , mapping = aes(x = year_week, y = gram_price)) + geom_boxplot()

dataset.price.week.new = ddply(dataset.price, c("year_week","invtype"), mutate,
                        five_percent = quantile(gram_price, 0.05),
                        ninty_five_percent = quantile(gram_price, 0.95))

dataset.price.week.new.clean = subset(dataset.price.week.new, 
                              gram_price > five_percent & gram_price < ninty_five_percent)

price.week.new = ddply(dataset.price.week.new.clean, 
                   c("year_week","invtype"),
                   summarise,
                   gram_price_ave = mean(gram_price),
                   gram_price_max = max(gram_price),
                   gram_price_min = min(gram_price),
                   gram_price_median = median(gram_price),
                   gram_price_std = sd(gram_price))

inhalation2 = subset(price.week.new, invtype == "Marijuana Extract for Inhalation ")
infused2 = subset(price.week.new, invtype== "Marijuana Mix Infused ")
packaged2 = subset(price.week.new, invtype == "Marijuana Mix Packaged ")
usable2 = subset(price.week.new, invtype == "Usable Marijuana ")

# the distribution of average gram_price of each inventory type in each week
q1= ggplot(data = inhalation2, mapping = aes(x= as.factor(year_week), y = gram_price_ave)) + 
  geom_point() +
  labs(title = 'average gram_price by weeks \n Marijuana Extract for Inhalation(5%-95%)' )+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))

q2 = ggplot(data = infused2, mapping = aes(x= as.factor(year_week), y = gram_price_ave)) + 
  geom_point() +
  labs(title = 'average gram_price by weeks \n Marijuana Mix Infused(5%-95%)' )+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
 
q3 =  ggplot(data = packaged2, mapping = aes(x= as.factor(year_week), y = gram_price_ave)) + 
  geom_point() +
  labs(title = 'average gram_price by weeks \n Marijuana Mix Packaged(5%-95%)' )+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))

q4 =  ggplot(data = usable2, mapping = aes(x= as.factor(year_week), y = gram_price_ave)) + 
  geom_point() +
  labs(title = 'average gram_price by weeks \n Usable Marijuana(5%-95%)' )+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))

gridExtra::grid.arrange(q1, q2, q3, q4,ncol=2)

ggplot(data = price.week.new, mapping = aes(x = as.factor(year_week), y = gram_price_ave, color = invtype)) + 
  geom_point() +
  labs(title = 'average gram_price by weeks \n different invtype (5%-95%)' ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
```


#### distribution of avegram_price of each inventory type in each month

```{r}
# calculate the average gram_price of each inventory type in each month
price.month = ddply(dataset.price, 
                   c("year_mo","invtype"),
                   summarise,
                   gram_price_ave = mean(gram_price),
                   gram_price_max = max(gram_price),
                   gram_price_min = min(gram_price),
                   gram_price_median = median(gram_price),
                   gram_price_std = sd(gram_price))

# the distribution of average gram_price of each inventory type in each month
ggplot(data = price.month, mapping = aes(x = year_mo, y = gram_price_ave, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'average gram_price by months and inventory type')+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))

# the distribution of median gram_price of each inventory type in each month
ggplot(data = price.month, mapping = aes(x=year_mo, y = gram_price_median, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'median gram_price by months and inventory type')+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
```

### Analysis on Price Using the second method: total revenue/total amount

#### distribution of total revenue/total amount of each inventory type in each week
```{r calculate price}
# calculate the average price of each week each inventory type
price.week2 = ddply(dataset.price, c("year_week","invtype"),
                    summarise, 
                    tprice = sum(price_adjust), 
                    tusableweight = sum(usableweight_adjust))
price.week2 = mutate(price.week2, 
                     average_price = tprice/tusableweight) 

# distribution of average price for each week each inventory type
ggplot(data = price.week2, mapping = aes(x = year_week, y = average_price, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'average price by week and inventory type')+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
```


#### distribution of total revenue/total amount of each inventory type in each week
```{r calculate price2}
# calculate the average price of each week each inventory type
price.month2 = ddply(dataset.price, c("year_mo","invtype"),
                    summarise, 
                    tprice = sum(price_adjust), 
                    tusableweight = sum(usableweight_adjust))
price.month2 = mutate(price.month2, 
                     average_price = tprice/tusableweight) 

# distribution of average price for each week each inventory type
ggplot(data = price.month2, mapping = aes(x = year_mo, y = average_price, group = invtype, color = invtype)) + 
  geom_line() + 
  labs(title = 'average price by months and inventory type')+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 7))
```